%% -*- mode: latex; mode: outline-minor; fill-column: 108 -*-
%% Title:  grammar
%%
%% $HeadURL: http://rads-svn.sce.carleton.ca:8080/svn/lqn/trunk-V5/doc/userman/infile.tex $
%% Original Author:     Greg Franks <greg@sce.carleton.ca>
%% Created:             Tue Jul 18 2006
%%
%% ----------------------------------------------------------------------
%% $Id: infile.tex 10182 2011-03-25 14:55:52Z greg $
%% ----------------------------------------------------------------------

\documentclass{article}
\usepackage{times}
\usepackage{bnf}

\oddsidemargin  18pt
\evensidemargin  18pt
\textwidth   6.5in
%\topmargin   0pt
%\textheight  8.5in
%\marginparwidth 0pt
%\marginparsep 0pt
\setcounter{secnumdepth}{0}

\begin{document}

\title{SRVN Input File Format\\ {\small\rm\$Id: infile.tex 10182 2011-03-25 14:55:52Z greg $\$$}}
\author {Dorina C. Petriu \and R. Gregory Franks \and Alex Hubbard}
\date{\today} 
\maketitle

\section{SRVN Input File Format}
The Stochastic Rendezvous Network (SRVN) is a modelling method under
development, and is currently being upgraded and extendend. Due to
this fact, the input file format was chosen to be as flexible as
possible, and easy to be expand when the need for new input data
occurs.

The SRVN input file has four sections, each consisting of a variable
length list, possibly with other lists nested therein.  The {\em end
of list} marker is $-1$.

The input file may be annotated with comments.\marginpar{\chgflg{3}}
Comments start with the octothorpe symbol, `{\tt \#}', and finish at the
end of the line.  Comments may appear anywhere except within strings.

\begin{description}
\item [General information section] gives input parameters related to
  the numerical computation.  It consists of five fields:
  \begin{enumerate}
  \item the convergence value,
  \item the underrelaxation coefficient,
  \item the maximum number of iterations,
  \item the number of iterations after which intermediate results
    are printed,
  \item and a field for {\em comments on model} which is not
    interpreted by the analysis program.
  \end{enumerate}
  
\item [Processor information section] declares the total number of
  processors in the system, and the corresponding scheduling policy for
  each processor.  The scheduling policy may be First Come First Served
  (FCFS) (the default value), Priority Preemptive Resume (PPR) or Head
  Of the Line (HOL)\footnote{Also known as Priority, Non-preemptive.}.
  The processors are named using symbolic identifiers.
  
\item [Task information section] declares the total number of tasks in
  the system, and the following parameters for each task: task type
  (reference or non-reference), the list of entries belonging to this
  task, the processor on which the task is allocated, and task priority
  (if it applies). The tasks are named using symbolic identifiers. Each
  task must have at least one entry because the workload parameters are
  associated with entries.  As a particular case, a reference task has
  one and only one entry, which is never called by other tasks
  (entries).
  
\item [Entry information section] declares the total number of entries
  and the workload parameters associated with each. The entries are
  named using symbolic identifiers. Each entry is associated to a task
  by way of an {\em entry list} for the task. (The model description is
  invalid if an entry is associated to more than one task, or no task at
  all).  In the BNF description below, the non-terminals {\em
  from\_entry\/} and {\em to\_entry\/} must belong to separate tasks.

The entry workload is divided in phases.  The input file format does
not limit the number of phases: it uses a variable length list with
parameters ordered by phase.  Default values are assumed for all
undeclared parameters.  The parameters declared for entries are:

\begin{itemize}
\item open arrival rate to entry (default = 0);
\item list of mean service times per phase (default = 0);
\item list of phase type flags (default = 0):
  \begin{enumerate}
  \item exponentially distributed service time and geometrically
    distributed number of calls;
  \item deterministic service time and deterministic number of
    calls;
  \item exponential service time and deterministic number of calls;
  \end{enumerate}
\item coefficient of variation of the service time slice between two
  calls, given by phase (default = 1.0);
\item mean number of RNVs per phase from this entry to other entries
  (default = 0.0);
\item mean number of nonblocking sends per phase from this entry to
  other entries (default = 0.0).
\item probability of forwarding a rendezvous to another entry (default
  = 0.0).
\item internal messaging between entries.
\end{itemize}
\end{description}

\section{BNF Description of SRVN Input File Format}

\input{input-grammar}

\subsection{Other Solvers}

\subsubsection{Lqsim}
The SRVN input file is also used by the layered queueing network simulator.
\marginpar{\chgflg{3}} Certain optional nonterminals only apply to the
simulator (for example, {\em cpu\_quantum}).  These nonterminals are
identified by a `\dag'.  If they are present as input to the SRVN
solver, they are ignored.  Mandatory fields marked by a `\ddag' are
ignored by the simulator.

\subsubsection{PetriSRVN}

Mandatory fields marked by a `\ddag' are ignored by the Petri net
solver.

\subsubsection{Layered Queueing Network Solver}


\subsubsection{Bounds Solver}

\end{document}
